---
title: "TP3"
author: "A REMPLIR ABSOLUMENT"
output: 
  html_document:
    number_sections: false
    toc_float: true
---

```{r setup, include=FALSE, message=FALSE}
#see full list >knitr::opts_chunk$get()
knitr::opts_chunk$set(echo = TRUE, fig.align="center", prompt = TRUE, comment="",
                      message = FALSE, warning=FALSE)
question <- 1

```


# Data preparation 


## Question 1



```{r, echo=FALSE}
# lecture du data
BreastCancer <- read.csv("BreastCancer.csv", header=TRUE, stringsAsFactors = TRUE) 

# Exploration :
str(BreastCancer)
head(BreastCancer)

# Clean data :
table(is.na(BreastCancer))
BreastCancer <- na.omit(BreastCancer)
table(is.na(BreastCancer))

```


## Question 2 : ?? 

```{r}
head(BreastCancer)
explvar <- c("Clump.Thickness", 
             "Uniformity.of.Cell.Size",
             "Uniformity.of.Cell.Shape",
             "Marginal.Adhesion",
             "Single.Epithelial.Cell.Size",
             "Bare.Nuclei",
             "Bland.Chromatin",
             "Normal.Nucleoli",
             "Mitoses")
summary(BreastCancer[, explvar])
summary(log(BreastCancer[explvar]))


```

Make (advanced) bivariate scatterplots which reveals some affine patterns.
```{r, echo=FALSE}
#make 2D-density
panel.density <- function(x, y, nbgrid = 100, nbcol = 100, 
                          col4ramp = c("green", "yellow", "orange", "red"), ...) 
{
  require(MASS)
  k <- try(kde2d(x, y, n = nbgrid), silent = TRUE)
  if(!inherits(k, "try-error"))
  image(k, col = colorRampPalette(col4ramp)(nbcol), add = TRUE, 
    xlim = range(x, na.rm = TRUE), ylim = range(y, na.rm = TRUE))
}
#make histogram
panel.hist <- function(x, ...) 
{
    usr <- par("usr")
    on.exit(par(usr))
    par("usr" = c(usr[1:2], 0, 1.5))
    his <- hist(x, plot = FALSE)
    breaks <- his$breaks
    nB <- length(breaks)
    y <- his$counts
    y <- y/(1.1*max(y))
    rect(breaks[-nB], 0, breaks[-1], y, col = rgb(0, 1, 1, alpha = 0.5), ...)
}
pairs(BreastCancer[, c(explvar, "Class")], lower.panel=panel.density,
      diag.panel = panel.hist, cex=0.5)

logBreastCancer <- log(BreastCancer[ , explvar])
pairs(logBreastCancer, lower.panel=panel.density,
      diag.panel = panel.hist, cex=0.5)
```


```{r}
library(corrplot)
mat.cor <- cor(log(BreastCancer[ , explvar]))
corrplot(mat.cor)
```


## Question 3

# Principal component analysis using `stats` package



## Question 4

-  prcomp() : effectue une ACP en utilisant decompostion en valeurs singuliéres SVD

- princomp() : effectue une ACP en utilisant la decompistion en valeures propres de la matrice de covariance V


par défaut ce deux fonction effectue l'ACP Canonique ie Metrique M = I 


## Question 5

```{r}
X <- as.matrix(log(BreastCancer[, explvar]))
breast.prcomp <- prcomp(X, scale=FALSE)
breast.princomp <- princomp(X, cor=FALSE)
```


### Subquestion (a)

```{r}
X.cov <- cov(X)
dim(X.cov)

# on calcul les vecteures propres 
diag.V <- eigen(X.cov)
val.prop.V <- diag.V$values
vec.prop.V <- diag.V$vectors


# check taht lambda_k = s_k^2 
  
cbind("s_k^2" = breast.prcomp$sdev^2 , "lambda_k" =val.prop.V )

```


### Subquestion (b)

```{r}
# Check the first eigenvector and the first axis breast.prcomp$loadings[,1] :
cbind("u_1" = vec.prop.V[,1] , "princomp$loadings[,1]" = breast.princomp$loadings[,1])

```


### Subquestion (c)


```{r}
# la matrice d'observation ie C : la matrice de composnate principale 
# C = YMU or M = I d'ou C = YU 

Y <- scale(X ,center = TRUE , scale = FALSE)
U <- diag.V$vectors
C <- Y %*% U 

# On compare C avce breast.prcomp$x
C
breast.prcomp$x
```


## Question 6


### Subquestion (a)

```{r}
# Center the observations 
X.s <- scale(X, center = TRUE, scale = FALSE) /sqrt(nrow(X) - 1)

```



### Subquestion (b)

```{r}
# Using svd(), check the explained variances against the output breast.prcomp$sdev :

# decompostion en valeurs singuliéres : UΣV^T
decomp.svd <- svd(X.s)
val.prop.svd <- decomp.svd$d

# Comparasion :
cbind(val.prop.svd , breast.prcomp$sdev)

```



### Subquestion (c)

```{r}
# Using svd(), check the right singular vectors against the output breast.prcomp$rotation
decomp.svd$v
breast.prcomp$rotation

# Verification des QQ prop des composantes principales
round(mean(breast.prcomp$rotation[,1]))
round(mean(breast.prcomp$rotation[,1]*t(breast.prcomp$rotation[,1])))
 
```



## Question 7


```{r}
# Interpret the values of explained variances. Use summary() and plot() on breast.prcomp :
summary(breast.prcomp)

```


```{r}
# le comportement du plot avce un objet de type prcomp 

plot(breast.prcomp , main = "Répresentation du variances en fonction du k" ,
     xlab ="k" , col = "blue")


```


## Question 8


```{r}
biplot(breast.prcomp)
```


```{r}
plot(breast.prcomp$x[, c(1,2)], 
     col=1+1*(BreastCancer$Class == "malignant") , pch = 20 , 
     main = "Projetction des indvidus sur PC1 , PC2" , xlab = "PC1" , 
     ylab ="PC2" )
grid()
legend("topleft" , legend = c("bening" ,"malignant") , col = c(1 ,2) , pch = 20 )

```

# Principal component analysis using `FactoMineR` package



## Question  9

## Question 10

## Question 11

