---
title: "SOlution"
author: "Nebil"
output: 
  html_document:
    number_sections: false
    toc_float: true
---

# Data preparation 

```{r}
df <- mtcars 

head(df)
colnames(df)

library(dplyr)
```



# Q1 :

### Return rows of cars that have an mpg value greater than 20 and 6 cylinders

```{r}

df %>% filter(mpg > 20 , cyl == 6 )


```


# Q2 :

### Reorder the Data Frame by cyl first, then by descending wt.

```{r}
df %>% 
  arrange(cyl , desc(wt)) %>%  
  head()

```



# Q3 :

### Select the columns mpg and hp

```{r}

df %>% 
  select(mpg , hp) %>% 
  head()

```


# Q4 :

### Select the distinct values of the gear column

```{r}

df %>% distinct(gear)

unique(df$gear)

```


# Q5 :

### Create a new column called "Performance" which is calculated by hp divided by wt

```{r}

df %>% 
  mutate(Performance = hp/wt) %>% 
  head()


```

# Q6 :

### find the mean mpg value using dplyr

```{r}

df %>% summarise(mean_mpg = mean(mpg))

```


# Q7 :

### Use pipe operators to get the mean hp value for cars with 6 cylinders.


```{r}
df %>% 
  filter(cyl == 6) %>% 
  summarise(mean_hp = mean(hp))
# 122.2857	
```



















